---
title: "MongoDB exercise"
date: "29 November 2022"
output: html_document
---

In this document you will try to replicate with MongoDB what we did using SQLite last week. We will therefore use MongoDB through R as well. Should you work with MongoDB in the future, you might use it e.g. via the MongoDB Shell (command line) or MongoDB Compass (a graphical user interface).

Loading packages:

```{r}
library("tidyverse")
library("mongolite")
library("DBI")
library("microbenchmark")
library("stringi")
```

## Building the Mongo database

1. Create a database and collection. 

```{r, eval=FALSE}
# Creates an R object linked to the "congress_posts_2017" collection
collection <- mongo(collection = "congress_posts_2017", db = "facebook")
```

2. Read in the congress table:

```{r, eval = FALSE}
# read the congress data first
congress <- read.csv("data/congress-facebook-2017.csv", stringsAsFactors = FALSE)
congress <- congress %>% rename(role = type)
```

3. Loop over all the posts tables and merge each with the relevant information from the congress table. Then store the result as a document in your Mongo database:

```{r, eval=FALSE}

fls <- list.files("data/posts", full.names=TRUE)

for (f in fls){
  
  # Status update
  message(f)

  # Read file into memory
  posts <- read.csv(f, stringsAsFactors = FALSE)
  
  # Convert screen name and date
  posts$screen_name <- as.character(posts$screen_name)
  posts$datetime <- parse_datetime(posts$datetime)
  
  # Merge with congress table information
  posts <- congress %>% inner_join(posts, by = "screen_name") 
  posts$message <- stri_unescape_unicode(posts$message) # solve some encoding issues
  
  # Adding into MongoDB collection
  collection$insert(posts)
  
}

```


```{r}
# Return the number of documents
collection$count()

# dbGetQuery(dbsql, 'SELECT * FROM posts LIMIT 5')
collection$find('{}', limit=5)  # '{}' indicates everything

# Note: You can remove (all) documents with
#collection$remove('{}')

collection$disconnect()
```

## Querying the Mongo database

4. Connect to the SQLite database from last week:

```{r}
# Change path to where you stored the SQLite database after you created it
dbsql <- dbConnect(RSQLite::SQLite(), "data/facebook-db.sqlite")
```


```{r}
# Connecting to the collection within the facebook database
collection <- mongo(collection = "congress_posts_2017", db = "facebook")

# Returning the first five documents
collection$find(query = '{}', limit=5) # '{}' indicates everything

# Returning the first five records from the SQL database (would need to join for equivalent)
dbGetQuery(dbsql, 'SELECT * FROM congress LIMIT 5')
```

5. Of your own choosing show examples of __SELECT__:

```{r}
# Querying just one column
collection$find(query = '{}', fields = '{"name": true}', limit = 10)

dbGetQuery(dbsql, "SELECT name FROM congress LIMIT 10")
```


6. Of your own choosing show examples of AND operator:

```{r}
collection$find(query = '{"type": {"$ne": "photo"}, "likes_count": {"$gt": 500}}', 
        fields = '{"from_name": true, "type": true, "date": true, "likes_count": true}', 
        limit = 10)

# Longer version with explicit naming of and operation
#collection$find(query = '{"$and": [{"type": {"$ne": "photo"}}, {"likes_count": {"$gt": 500}}]}', 
#        fields = '{"from_name": true, "type": true, "date": true, "likes_count": true}', 
#        limit = 10)

dbGetQuery(dbsql, "SELECT from_name, type, date, likes_count 
           FROM posts
           WHERE type != 'photo' AND likes_count > 500
           LIMIT 10")
```

7. Of your own choosing show examples of OR operator:

```{r}
collection$find(query = '{"$or": [{"type": "photo"}, {"type": "video"}]}', 
        fields = '{"from_name": true, "type": true, "date": true, "likes_count": true}', 
        limit = 10)

dbGetQuery(dbsql, "SELECT from_name, type, date, likes_count 
           FROM posts
           WHERE  type = 'photo' OR type = 'video'
           LIMIT 10")
```

8. Of your own choosing show examples of Membership / IN:

```{r}
collection$find(query = '{"type": {"$in": ["photo",  "video"]}}', 
        fields = '{"from_name": true, "type": true, "date": true, "comments_count": true}', 
        limit = 10)
dbGetQuery(dbsql, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE type IN ('photo', 'video')
           LIMIT 10")
```


```{r}
collection$find(query = '{"date": {"$regex": "2017-01-.{2}", "$options": "i"}}', 
        fields = '{"from_name": true, "type": true, "date": true, "likes_count": true}', 
        limit = 10)
dbGetQuery(dbsql, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE date LIKE '2017-01-__'
           LIMIT 10")


collection$find(query = '{"date": {"$regex": "2017-01.+", "$options": "i"}}', 
        fields = '{"from_name": true, "type": true, "date": true, "likes_count": true}', 
        limit = 10)
dbGetQuery(dbsql, "SELECT from_name, type, date, comments_count 
           FROM posts
           WHERE date LIKE '2017-01%'
           LIMIT 10")


collection$find(query = '{"message": {"$regex": "london", "$options": "i"}}', 
        fields = '{"from_name": true, "type": true, "date": true, "likes_count": true, "message": true}', 
        limit = 10)
dbGetQuery(dbsql, "SELECT from_name, message, date
           FROM posts
           WHERE message LIKE '%london%'
           LIMIT 10")
```

9. Select rows/documents after a certain date (However, this is much more convenient in the SQLite implementation).

```{r}
dbGetQuery(dbsql, "SELECT from_name, type, date
           FROM posts
           WHERE date > '2017-01-01'
           LIMIT 10")

# Specifying the date is a bit too complicated, needs to convert the date into milliseconds
d <- as.integer(as.POSIXct(strptime("2017-01-01", "%Y-%m-%d"))) * 1000
collection$find(query = paste0('{"datetime": {"$gt": {"$date": {"$numberLong": "', d, '" } } } }'), 
        fields = '{"from_name": true, "type": true, "date": true}', 
        limit = 10)
```

10. Show examples of `*$aggregate()`. 

```{r}
collection$aggregate('[ {"$group": {"_id": "$from_name", "count": {"$sum": 1}}},
                  {"$limit": 10} ]')

dbGetQuery(dbsql, 
  "SELECT from_name, COUNT(*) AS post_count
  FROM posts
  GROUP BY from_name
  LIMIT 10")


# Conditional aggregate (only Republican politicians)
collection$aggregate('[{ "$match": {"party": "Republican"}}, 
                  {"$group": {"_id": "$from_name", "count": {"$sum": 1}}},
                  {"$limit": 10}]')
```

10. Show examples of `"$sort"` after find or aggregate.

```{r}
# Sort by type_count
collection$aggregate('[{"$group": {"_id": "$type", "type_count": {"$sum": 1}}},
                  {"$sort": {"type_count": 1}}]')
dbGetQuery(dbsql, 
  "SELECT type, COUNT(type) AS type_count
  FROM posts
  GROUP BY type
  ORDER BY type_count")


# And in descending orders
collection$aggregate('[{"$group": {"_id": "$type", "type_count": {"$sum": 1}}},
                  {"$sort": {"type_count": -1}}]')
dbGetQuery(dbsql, 
  "SELECT type, COUNT(type) AS type_count
  FROM posts
  GROUP BY type
  ORDER BY type_count DESC")
```

11. Find most popular post:

```{r}
collection$find(query = '{}',
             field = '{"from_name": true, "message": true, "likes_count": true, "datetime": true}',
             sort = '{"likes_count": -1}',
             limit = 1)

dbGetQuery(dbsql, 
  "SELECT from_name, message, likes_count, datetime
  FROM posts
  ORDER BY likes_count DESC
  LIMIT 1")
```

12. Find highest comment to like ratio:

```{r}
# We subset only posts with 1000 likes or more to avoid outliers

collection$aggregate('[{"$match" : {"likes_count": {"$gt": 1000}}},
                  {"$project": {"from_name": true, "message": true, "likes_count": true, "comments_count": true, "date": true,
                  "comment_like_ratio": {"$divide": ["$comments_count", "$likes_count"]}}},
                  {"$sort": {"comment_like_ratio": -1}},
                  {"$limit": 5}]') 

dbGetQuery(dbsql,
  "SELECT from_name, message, likes_count, comments_count, date,   
      comments_count/likes_count AS comment_like_ratio
  FROM posts
  WHERE likes_count > 1000
  ORDER BY comment_like_ratio DESC
  LIMIT 5")
```

## Joins in MongoDB

Joining in MongoDB is possible, however, easier in relational database which are built for this process. 

13. Creating additional data:

```{r}
set.seed(123)
content <- as_tibble(congress$screen_name)
colnames(content) <- "screen_name"
content$some_data <- runif(n = nrow(content))
head(content)
```

14. Add the data to a new collection (so you would have two collections in the `facebook` database: `congress_posts_2017` and `additional_data`):

```{r}
another_collection <- mongo(collection = "additional_data", db = "facebook")
another_collection$insert(content)
another_collection$find(query = '{}', sort = '{"screen_name": -1}',limit=5) 
```

15. Merge this `additional_data` collection into your base collection `congress_posts_2017` via the `$lookup`. 

```{r}
join_output <- collection$aggregate('[
    { "$project": {"screen_name": true, "type": true}},
    { "$sort": { "screen_name": -1 } },
    { "$limit": 10 },
    { "$lookup": {
      "localField": "screen_name",
      "from": "additional_data",
      "foreignField": "screen_name",
      "as": "additional_data"
    } }]')
join_output
```


## Performance?

You should not done any tuning (e.g. indexing). 
16. Compare which is faster just for fun introducing the microbenchmark package:

```{r}
microbenchmark(sqlite = 
  dbGetQuery(dbsql, "SELECT from_name, type, date, likes_count 
           FROM posts
           WHERE type != 'photo' 
              AND likes_count > 500
           LIMIT 10"),
mongo = collection$find(query = '{"type": {"$ne": "photo"}, "likes_count": {"$gt": 500}}', 
        fields = '{"from_name": true, "type": true, "date": true, "likes_count": true}', 
        limit = 10), times = 200)
```


```{r}
microbenchmark(sqlite = 
  dbGetQuery(dbsql,
    "SELECT from_name, message, likes_count, comments_count, date,   
        comments_count/likes_count AS comment_like_ratio
    FROM posts
    WHERE likes_count > 1000
    ORDER BY comment_like_ratio DESC
    LIMIT 5"),
  mongo = collection$aggregate('[{ "$match" : {"likes_count": {"$gt": 1000}}},
                  {"$project": {"from_name": true, "message": true, "likes_count": true, "comments_count": true, "date": true,
                  "comment_like_ratio": {"$divide": ["$comments_count", "$likes_count"]}}},
                  {"$sort": {"comment_like_ratio": -1}},
                  {"$limit": 5}]'),
times = 50)
```


